Demand Forecasting of individual Probability Density Functions with Machine Learning

### 2020-08-30 LaTex/paper.tex
cosmetics, larger axis labels and legends, added authors

\author[1]{Felix Wick \thanks{felix.wick@blueyonder.com}}
\author[2]{Ulrich Kerzel \thanks{u.kerzel@iubh-fernstudium.de}}
\author[1]{Trapti Singhal \thanks{trapti.singhal@blueyonder.com}}
\author[1]{Martin Hahn \thanks{martin.hahn@blueyonder.com}}
\affil[1]{\small Blue Yonder GmbH (Karlsruhe, Germany)}
\affil[2]{\small IUBH Internationale Hochschule (Erfurt, Germany)}

@@

Demand forecasting is one of the main challenges for retailers and is at the core of supply chain business operations. Due to its stochastic nature, demand is difficult to forecast: It depends on many influencing factors and can be interpreted as a random variable that is described by an appropriate probability density function (PDF). Demand estimation is further complicated by the fact that retailers typically only observe realized sales rather than the actual demand, and in case the demand exceeds the current stock level the data become censored. Demand forecasting is an application of time series forecasting and it is important to note that demand as a random variable is not identically and independently distributed (i.i.d.) between different items and locations. While the probability distribution describing the demand can be attributed to a given family or parameterization, the exact parameters vary: Seasonal effects, finite life cycles of products, and the introduction of new products influence the demand distribution, as well as the local weather at the retail location or the retail location itself in terms of size, assortment range, customer diversity, and other factors. The retailers themselves also actively influence demand by using advertisements to highlight products, offering rebates or discounts for specific products as well as pursuing an active pricing strategy. This means that while we can generally assume that demand follows a specific type of probability distribution, its parameters are unique to the instance for which an estimate is required. For example, the probability distribution governing the demand of a particular item is specific to the item, date, and retail location for which the forecast is made, and depends on a wide range of further influencing factors.

Thanks to the availability of Big Data today, a vast amount of these data associated with demand, such as historic sales records, information about promotional events or advertisements, pricing information, local weather at retail locations, seasonal information, as well as many further variables, can be captured, stored, and processed. Different methods, such as traditional univariate time series forecasting or modern supervised machine learning algorithms, can then be used to predict the demand distribution for individual products sold at specific locations, corrected for censored data. Demand forecasting for retailers often requires to predict millions of item-store combinations daily, and traditional univariate time series forecasting methods operate on each of these many individual time series separately. Machine learning, on the contrary, optimizes all individual time series together, improving the generalizability of the method by exploiting common aspects between them. This leads to a drastic reduction in variance of the model, and in turn to an improvement of the forecast quality of each of the individual time series. Furthermore, machine learning enables a natural way to include the many exogenous variables influencing demand as features of the model, what reduces also the bias compared to traditional time series forecasting approaches which mainly rely on the historic sales.

In order to make operational decisions, an optimal demand point estimator has to be defined, that can be used to derive ordering decisions in the replenishment process of the retailer. The ordering decision process is complicated by a range of factors: Even in the case of perfect demand forecasts, the decision maker has to consider lot-sizes defined by the wholesaler or manufacturer as well as to balance conflicting metrics to reach an optimal decision. Ordering too few items may result in out of stock situations leading to unrealized demand and unsatisfied customers. Ordering too many items results in excess inventory, which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoilt items need to be disposed of at additional cost and potentially even environmental impact. This situation is particularly noticeable in the so called ``ultra-fresh'' category, which includes items such as bakery products, ready-meals, fresh diary products, or certain meat products such as ground meat. These items typically have a shelf-life ranging from less than a business day to a few business days at most, with a continuous spectrum in between, depending on the exact item. In many situations, additional constraints have to be considered to reach an optimal ordering decision: Delivery cycles of items may vary depending on the type of item and the wholesaler or manufacturer from which they are procured. Retailers also operate at a given service level to guarantee that a certain level of demand can be fulfilled. The exact service level typically depends on the overall business policy of the retailer and may also depend on individual products, ranging from ``never-out-of-stock'' items to a service level exceeding, e.g., 90\%. So, demand forecasts are ideally made in the form of full individual probability density distributions, from which optimal point estimators can be derived, which in turn serve as starting point to determine the subsequent ordering decisions.

@@

The direct approach is often referred to as ``data-driven newsvendor'' and discussed e.g. in \cite{beutel2012safety,ban2019big,bertsimas2020predictive, oroojlooyjadid2020applying}. It aims to avoid estimating the underlying probability distribution for demand and use the available data (historic sales records and further variables) to derive the operational decisions (i.e., the order quantity) directly. An overview of a range of different approaches can also be found in \cite{huber2019data}. Although the integrated approach seems preferable at first glance, since it avoids determining the full demand distribution and results directly in the desired operational decision, the indirect approach via demand forecasts offers some substantial advantages. First, demand forecasts in form of full PDFs can be used to simulate the performance on an individual level, and, for example, optimize the impact on business strategy decisions on conflicting metrics such as out of stock (i.e., lost sales) and waste rate. From a practitioners perspective, separating the demand forecast from the operational decisions (i.e., calculating the order quantities for the next delivery cycle) enables longer-term planning and reduces the complexity, as it avoids coupling delivery schedules of multiple wholesalers and manufacturers with the forecast of customer demand. It also allows to share long-term demand predictions with other business units or external vendors and wholesalers to ease their planning for the production and supply chain processes upstream of the retailer. From the perspective of the industrial practice of a supplier of supply chain methods and tools, modeling the demand separately from deriving the subsequent orders has the additional benefit that multiple retail chains can benefit from any improvement in the model description, even if the specific retailers are unrelated to each other. Additionally, a purely data-driven approach going from the observed data directly to the operational decision (such as the order quantity) does not allow to analyze the data-generating process, i.e., the mechanism behind the stochastic behavior of the customer demand. This is crucial, for example, if a causal analysis is planned, such as a study of the effect of promotions, advertisements, price changes, or other demand shaping factors in either Pearl's do-calculus \cite{PearlCausality} or Rubin's potential outcomes framework \cite{rubin1974estimating}. Using an operational quantity such as the order quantity will in most cases act as an insufficient proxy for the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that one may not be able to  fully controlled for.

Therefore, the main objective in solving the inventory control problem is to determine the underlying demand distribution. The simplest approach is to use the observed sales records and forecast these via sample average approximation (SAA), see e.g. \cite{shapiro2014} for an overview, or as a time series (see e.g. \cite{alwan2016}), for example by means of autoregressive integrated moving average (ARIMA) \cite{BoxJenkins}, exponential smoothing \cite{ExponentialSmoothing}, or structural time series models \cite{StructuralTS}. However, these approaches do not make use of any data apart from the sales records themselves, although we know that many variables such as price, advertisements, or weather influence, and others are highly correlated with the demand. Saghafian and Tomlin \cite{saghafian2016newsvendor} propose to include partial information about the distribution in the derivation of the operational decision, i.e., the calculation of the optimal order quantity. As described in sec. \ref{sec:intro}, modern supervised machine learning techniques can overcome many of the issues of the more traditional approaches and therefore are the method of choice in most use cases today (see e.g. \cite{tsforecasting,ZHANG199835,rnn,transformer}).

In order to be able to fully optimize the operational decision, it is critical to reconstruct a full demand distribution. This also implies that a simple point-estimator, as provided by the most common statistical techniques and machine learning approaches, will not suffice. Instead, we need to determine the full demand distribution from data, conditional on the relevant variables such as date, location, and item, taking all auxiliary data such as article characteristics, pricing, advertisements, retail location details, etc. into account. This can be done in several ways: Quantile regression \cite{koenker2001, wen2017} can be implemented in various frameworks and used to estimate a range of quantiles for each predicted distribution, from which an empirical probability distribution can be interpolated. Using a dedicated neural network \cite{Feindt2006190}, either the full PDF or a defined range of quantiles can be calculated directly from the data for each individual prediction without assuming an underlying model. Alternatively, one can assume a given demand model and fit the model parameters instead of reconstructing the complete distribution \cite{astonpr373, SALINAS20201181}. This approach is computationally favorable and usually more robust, as fewer parameters need to be estimated. Empirically, one can determine the best fitting distribution from data \cite{adan1995}. However, given the stochastic nature of the demand, such an empirically determined distribution is not expected to be stable and prone to sudden changes. Instead, the choice of the demand distribution should be motivated by theoretic considerations. The discrete demand is typically modelled as a negative binomial distribution (NBD), also known as Gamma-Poisson distribution \cite{Ehrenberg1959,Ehrenberg1967,Ehrenberg1972,Chatfield1973,Schmittlein_1985}. This distribution arises if the Poisson parameter $\mu$ is a random variable itself, which follows a Gamma distribution. The NBD has two parameters, $\mu$ and $ \sigma^2 > \mu$, and is over-dispersed compared to the Poisson distribution, for which $\mu = \sigma^2$. Hence, for each ordering decision, the model parameters $\mu$ and $\sigma$ need to be determined for each item at the required granularity, typically for each sales location and ordering time, depending on all auxiliary data describing article details, retail location, and influencing factors such as pricing and advertisement information.

@@

This work demonstrates how the explainable machine learning algorithm Cyclic Boosting \cite{Wick2019} can be used to model the demand distribution at the granularity needed by the retailer. Typically, this means that the full demand PDF has to be estimated per product for each sales location and opening day, conditional on a wide range of variables such as weather, prices, promotions, etc. In contrast to using a ``black-box'' machine learning model, Cyclic Boosting allows to interpret each individual prediction in terms of influence of the different features.

@@

To predict an individual PDF using a parametric approach, one has to rely on a model assumption about the underlying distribution of the random variable to be predicted. As discussed earlier, the NBD is well routed in theoretical arguments to model customer demand. Its parameters can be modeled by two independent models, one to estimate the mean and the other for the variance. At least in principle, any method can be used. However, as discussed in sec. \ref{sec:LitRev}, machine learning algorithms are ideally suited for the task of demand forecasting and in the following, we will use the Cyclic Boosting algorithm to benefit in particular from explainable decisions rather than black-box approaches. Furthermore, the regularization approach used during training of the Cyclic Boosting algorithm allows a dedicated treatment of the underlying NBD model, which is another major benefit compared to a standard ``off-the-shelf'' machine learning algorithm. This means we use two subsequent Cyclic Boosting models, in order to estimate the parameters of each individual PDF that we need to forecast, where the first model is used to estimate the mean and the second to estimate the variance. Hereby, the features may or may not differ between the mean and variance estimation models, and it can be beneficial to include the corresponding mean predictions as feature in the variance model. The assigned mean and variance predictions can then be used to generate individual PDFs using the parameterization of the NBD for each sample.

@@

Cyclic Boosting is a type of generalized additive model using a cyclic coordinate descent optimization and featuring a boosting-like update of parameters. Major benefits of Cyclic Boosting are its accuracy, performance, even at large scale, and providing fully explainable predictions, which are of vital importance in practical applications.

The main idea of this algorithm is the following: First, each feature, denoted by index $j$, is discretized appropriately into $k$ bins to reflect the specific behaviour of the feature. The global mean $\mu$ is determined from all values $y$ of the  target variable $Y \in [0,\infty)$ observed in the data. Single data records, for example, the sales of a specific product-location-date combination along with all relevant features, are indexed by $i$.
The individual predictions $\hat{y_i}$  can then be calculated as:
\begin{equation} \label{eqn:cb}
\hat{y}_i = \mu \cdot \prod \limits_{j=1}^p f^k_j \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
\end{equation}
The factors $f^k_j$ are the model parameters that are determined iteratively from the features until the algorithm converges. During training, regularization techniques are applied to avoid overfitting and improve the generalization ability of the algorithm. The deviation of each factor from $f^k_j=1$ can then be used to explain how a specific feature contributes to each individual prediction.

@@

For a first qualitative assessment, we make use of the probability integral transform, see e.g. \cite{Angus1994,casella2002statistical}, which states that a random variable distributed according to the CDF of another random variable is uniformly distributed. So, in case of a correctly calibrated PDF prediction, the distribution of the corresponding actually observed CDF values of the individual PDF predictions is expected to be uniform, {\em regardless} of the shape of the predicted distribution, and any deviation can be interpreted as a hint that the predicted PDF is not fully correct \cite{diebold1998vevaluating}.

The CDF of a PDF $f(x)$ is defined as:

@@

In case of discrete probability functions, such as the negative binomial function, the same argument still holds, but the definition of the inverse quantile function is replaced by the generalized inverse: $F^{-1}(y) = \mathrm{\inf \left \{x : F(x)>y\right  \} }$ for $y \in [0,1]$, see e.g. \cite[p. 54]{casella2002statistical}. In order to obtain a uniform distribution for discrete PDFs that is comparable to the case of continuous distributions, the histogram holding the values of the CDF is filled using random numbers according to the intervals of the CDF. For example, if the sales of zero items accounts for 75\% of the observed sales distribution for this item, the value of the CDF function that is used to fill the histogram in case of zero actual sales is randomly chosen in the interval $[0, 0.75]$. Proceeding similarly for all other observed values, with the intervals from which to randomly choose values to fill in the histogram defined by the CDF values of the corresponding discrete sales value and the one below (e.g. for 3 actual sales: random pick between discrete CDF values for 2 and 3), the resulting histogram of CDF values should again be uniform as in the case of a continuous PDF.

@@

Then we compare the actual observed sales value (corrected for censored data if necessary) to different quantiles of the corresponding predicted distribution for each data record and average over a larger data sample. For example, if we wanted to check that the median of the distribution, corresponding to the $0.5$ quantile, is predicted correctly by the machine learning model, we would compare the quantile value $0.5$ to the ratio of quantile values (again randomly chosen from the corresponding range of CDF values for discrete target values, as described in sec. \ref{sec:cdf_histo}) of observed sales records being lower/higher than $0.5$. In other words, in case of the median, 50\% of the ex post observed target values should be observed below the median of the corresponding individual predicted PDF and 50\% above.

In order to judge whether the overall shape of the predicted distributions is predicted correctly, we repeat this procedure for a range of quantiles, for example $q = 0.1, 0.2, \ldots, 0.9$. However, we are free to choose which quantiles to look at, and in specific situations it might be advisable to look at the tails of the distribution in more detail, to make sure that even relatively rare events are estimated correctly by the machine learning algorithm, and add more quantiles for comparison in the region between, say $q = 0.95$ and $q = 0.99$. In the following, we call this method {\em quantile profile plot}.

@@

In the same way as the method of filling all observed CDF values of the individual PDF predictions in a histogram and compare to a uniform distribution (described in sec. \ref{sec:cdf_histo}), quantile profile plots do not work on individual PDF predictions but require a statistical population. However, calculating the fractions of the quantile profile plots globally, i.e., over all samples in the data set, might not reveal certain deficiencies for a subset, e.g., a specific store or product group in the example of retail demand forecasts. Therefore, we combine the approach described above with the method of profile plots, where the quantity on the $x$-axis of the quantile profile plot can be any variable of the data set at hand. Profile plots are akin to scatter plots and described in more detail in appendix \ref{sec:profile}.

@@

\noindent
The different \texttt{item\_store} values for mean (higher than $1$) and variance model factors (lower than $1$), including the static one-dimensional features \texttt{store\_id} and \texttt{item\_id} and their two-dimensional combination, represent the fact that this product-location combination sells more than the average of all product-location combinations, and the dispersion parameter $1/r$ in \eqn \eqref{eqn:variance_r} in turn tends towards lower values. The explanation of the model for the peak in sales and mean prediction on February 7, 2016 is the event Super Bowl. The variance prediction for this day is mainly made up from two competing factors, the Super Bowl event driving higher dispersion parameter and the high mean prediction driving lower dispersion parameter.

@@

\fig \ref{fig:invquant_mean} shows the quantile profile plot using the predicted mean as the $x$-axis of the graph, meaning that the distributions are grouped such that the 15 columns correspond to mean predictions in 12 intervals with widths of 5 from 0.0 to 60.0 and 3 remaining intervals $[60.0, 70.0]$, $(70.0, 80.0]$, and $(80.0, 100.0]$ (with mean predictions higher than 100.0 included in the highest interval), while we aggregate over all locations, items, and sales dates. The used quantile values are the same as in \fig \ref{fig:invquant_dayofweek}. While the Poisson model again shows a significant pattern of too narrow PDF predictions across the full range of mean prediction values, we can also see several deviations from the expected uniform behavior for the negative binomial model. For mean predictions around $30$, the negative binomial model deviates significantly from the expected behaviour, with the shape of the deviations pointing towards too broad PDF predictions (as also observed in the CDF histogram above). And for mean predictions higher than $80$, the pattern of the deviations indicate an underprediction of the mean parameter. The deviations at very low predicted mean values reflect the complications of slow-sellers described in the context of CDF histograms above.

@@

Evaluating these PDF predictions becomes the critical question. Common metrics, such as the mean absolute deviation or mean squared error, only take point estimates into account and are generally not suitable for the evaluation of predicted probability distributions. Furthermore, since the distributions are predicted for individual events, statistical techniques aiming at the comparison of distributions using many events cannot be used in general. Instead, novel techniques using both the histogram of the probability integral transformation as well as quantile profile histograms allow a detailed investigation into the behavior of the predictions. Finally, a quantitative assessment resulting in a single number can be obtained using metrics such as the earth mover distance in a comparison between the histogram of the probability integral transformation and the expected uniform distribution.

@@

In many cases, scatter plots are used to study the behaviour of two distributions or sets of data points visually. However, even for moderate amount of data, this approach becomes quickly difficult. To illustrate this, a sample of $(x,y)$ data points was obtained in the following way: The distribution of $x$ values was obtained by generating 5,000 samples of Gaussian distributed random numbers $X \sim {\cal N}(0.0,2.0)$ and the $y$ values are obtained via $Y \sim X +  {\cal N}(2.0,1.5)$. \fig \ref{fig:scatter} shows the marginal distributions for $x$ and $y$ as well as a scatter plot of $x$ vs. $y$.

>>>>

\author[1]{Felix Wick \thanks{felix.wick@blueyonder.com}}
\author[3]{Ulrich Kerzel \thanks{u.kerzel@iubh-fernstudium.de}}
\author[1]{Martin Hahn \thanks{martin.hahn@blueyonder.com}}
\author[1]{Moritz Wolf \thanks{martin.hahn@blueyonder.com}}
\author[2]{Trapti Singhal \thanks{trapti.singhal@blueyonder.com}}
\author[1]{Michael Feindt \thanks{michael.feindt@blueyonder.com}}
\affil[1]{\small Blue Yonder GmbH (Karlsruhe, Germany)}
\affil[2]{\small Blue Yonder India Private Limited (Bengaluru, India)}
\affil[3]{\small IUBH Internationale Hochschule (Erfurt, Germany)}

@@

Demand forecasting is one of the main challenges for retailers and is at the core of supply chain business operations. Due to its stochastic nature, demand is difficult to forecast: It depends on many influencing factors and can be interpreted as a random variable that is described by an appropriate probability density function (PDF). Demand estimation is further complicated by the fact that retailers typically only observe realized sales rather than the actual demand, and, in case the demand exceeds the current stock level, the data become censored. Demand forecasting is an application of time series forecasting and it is important to note that demand as a random variable is not identically and independently distributed (i.i.d.) between different items and sales locations. While the probability distribution describing the demand can be attributed to a given family or parameterization, the exact parameters vary: Seasonal effects, finite life cycles of products, and the introduction of new products influence the demand distribution, as well as the local weather at the retail location or the retail location itself in terms of size, assortment range, customer diversity, and other factors. The retailers themselves also actively influence demand by using advertisements to highlight products, offering rebates or discounts for specific products, as well as pursuing an active pricing strategy. This means that while we can generally assume that demand follows a specific type of probability distribution, its parameters are unique to the instance for which an estimate is required. For example, the probability distribution governing the demand of a particular item is specific to the item, date, and retail location for which the forecast is made, and depends on a wide range of further influencing factors.

Because of the access to  Big Data, a vast amount of these data associated with demand, such as historic sales records, information about promotional events or advertisements, pricing information, local weather at retail locations, seasonal information, as well as many further variables, can be captured, stored, and processed. Different methods, such as traditional univariate time series forecasting or modern supervised machine learning algorithms, can then be used to predict the demand distribution for individual products sold at specific locations, corrected for censored data. Demand forecasting for retailers often requires to predict millions of item-store combinations daily, and traditional univariate time series forecasting methods operate on each of these many individual time series separately. Machine learning, on the contrary, optimizes all individual time series together, improving the generalizability of the method by exploiting common aspects between them. This leads to a drastic reduction in variance of the model, and in turn to an improvement of the forecast quality of each of the individual time series. Furthermore, machine learning enables a natural way to include the many exogenous variables influencing demand as features of the model, what reduces also the bias compared to traditional time series forecasting approaches which mainly rely on the historic sales.

In order to make operational decisions, an optimal demand point estimator has to be defined, that can be used to derive ordering decisions in the replenishment process of the retailer. The ordering decision process is complicated by a range of factors: Even in the case of perfect demand forecasts, the decision maker has to consider lot-sizes defined by the wholesaler or manufacturer as well as to balance conflicting metrics to reach an optimal decision. Ordering too few items may result in out of stock situations leading to unrealized demand and unsatisfied customers. Ordering too many items results in excess inventory, which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoiled items need to be disposed of at additional cost and potentially even environmental impact. This situation is particularly noticeable in the so called ``ultra-fresh'' category, which includes items such as bakery products, ready-meals, fresh diary products, or certain meat products such as ground meat. These items typically have a shelf-life ranging from less than a business day to a few business days at most, with a continuous spectrum in between, depending on the exact item. In many situations, additional constraints have to be considered to reach an optimal ordering decision: Delivery cycles of items may vary depending on the type of item and the wholesaler or manufacturer from which they are procured. Retailers also operate at a given service level to guarantee that a certain level of demand can be fulfilled. The exact service level typically depends on the overall business policy of the retailer and may also depend on individual products, ranging from ``never-out-of-stock'' items to a service level exceeding, e.g., 90 percent. Therefore, demand forecasts are ideally made in the form of full individual probability density distributions, from which optimal point estimators can be derived, which in turn serve as starting point to determine the subsequent ordering decisions.

@@

The direct approach is often referred to as ``data-driven newsvendor'' and discussed e.g. in \cite{beutel2012safety,ban2019big,bertsimas2020predictive, oroojlooyjadid2020applying}. It aims to avoid estimating the underlying probability distribution for demand and use the available data (historic sales records and further variables) to derive the operational decisions (i.e., the order quantity) directly. An overview of a range of different approaches can also be found in \cite{huber2019data}. Although the integrated approach seems preferable at first glance, since it avoids determining the full demand distribution and results directly in the desired operational decision, the indirect approach via demand forecasts offers some substantial advantages. First, demand forecasts in form of full PDFs can be used to simulate the performance of the relevant metrics on the level of individual item, and, for example, optimize the impact on business strategy decisions on conflicting metrics such as out of stock (i.e., lost sales) and waste rate. From a practitioners perspective, separating the demand forecast from the operational decisions (i.e., calculating the order quantities for the next delivery cycle) enables longer-term planning and reduces the complexity, as it avoids coupling delivery schedules of multiple wholesalers and manufacturers with the forecast of customer demand. It also allows to share long-term demand predictions with other business units or external vendors and wholesalers to ease their planning for the production and supply chain processes upstream of the retailer. From the perspective of the industrial practice of a vendor of supply chain methods and tools, modeling the demand separately from deriving the subsequent orders has the additional benefit that multiple retail chains can benefit from any improvement in the model description, even if the specific retailers are unrelated to each other. Additionally, a purely data-driven approach going from the observed data directly to the operational decision (such as the order quantity) does not allow to analyze the data-generating process, i.e., the mechanism behind the stochastic behavior of the customer demand. This is crucial, for example, if a causal analysis is planned, such as a study of the effect of promotions, advertisements, price changes, or other demand shaping factors in either Pearl's do-calculus \cite{PearlCausality} or Rubin's potential outcomes framework \cite{rubin1974estimating}. Using an operational quantity such as the order quantity will in most cases act as an insufficient proxy for the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that one may not be able to  fully controlled for.

Therefore, the main objective in solving the inventory control problem is to determine the underlying demand distribution. The simplest approach is to use the observed sales records and forecast these via sample average approximation (SAA), see, e.g., \cite{shapiro2014} for an overview, or as a time series (see, e.g., \cite{alwan2016}), for example by means of auto-regressive integrated moving average (ARIMA) \cite{BoxJenkins}, exponential smoothing \cite{ExponentialSmoothing}, or structural time series models \cite{StructuralTS}. However, these approaches do not make use of any data apart from the sales records themselves, although we know that many variables such as price, advertisements,  weather influence, and others are highly correlated with the demand. Saghafian and Tomlin \cite{saghafian2016newsvendor} propose to include partial information about the distribution in the derivation of the operational decision, i.e., the calculation of the optimal order quantity. As described in sec. \ref{sec:intro}, modern supervised machine learning techniques can overcome many of the issues of the more traditional approaches, and therefore are the method of choice in most use cases today (see, e.g., \cite{tsforecasting,ZHANG199835,rnn,transformer}).

In order to be able to fully optimize the operational decision, it is critical to reconstruct a full demand distribution. This also implies that a simple point-estimator, as provided by the most common statistical techniques and machine learning approaches, will not suffice. Instead, we need to determine the full demand distribution from data, conditional on the relevant variables such as date, location, and item, taking all auxiliary data such as article characteristics, pricing, advertisements, retail location details, etc. into account. This can be done in several ways: Quantile regression \cite{koenker2001, wen2017} can be implemented in various frameworks and used to estimate a range of quantiles for each predicted distribution, from which an empirical probability distribution can be interpolated. Using a dedicated neural network \cite{Feindt2006190}, either the full PDF or a defined range of quantiles can be calculated directly from the data for each individual prediction without assuming an underlying model. Alternatively, one can assume a given demand model and fit the model parameters instead of reconstructing the complete distribution \cite{astonpr373, SALINAS20201181}. This approach is computationally favorable and usually more robust, as fewer parameters need to be estimated. Empirically, one can determine the best fitting distribution from data \cite{adan1995}. However, given the stochastic nature of the demand, such an empirically determined distribution is not expected to be stable and prone to sudden changes. Instead, the choice of the demand distribution should be motivated by theoretic considerations. The discrete demand is typically modeled as a negative binomial distribution (NBD), also known as Gamma-Poisson distribution \cite{Ehrenberg1959,Ehrenberg1967,Ehrenberg1972,Chatfield1973,Schmittlein_1985}. This distribution arises if the Poisson parameter $\mu$ is a random variable itself, which follows a Gamma distribution. The NBD has two parameters, $\mu$ and $ \sigma^2 > \mu$, and is over-dispersed compared to the Poisson distribution, for which $\mu = \sigma^2$. Hence, for each ordering decision, the model parameters $\mu$ and $\sigma$ need to be determined for each item at the required granularity, typically for each sales location and ordering time, depending on all auxiliary data describing article details, retail location, and influencing factors such as pricing and advertisement information.

@@

This work demonstrates how the explainable machine learning algorithm Cyclic Boosting \cite{Wick2019} can be used to model the demand distribution at the granularity required by the retailer. Typically, this means that the full demand PDF has to be estimated per product for each sales location and opening day, conditional on a wide range of variables such as weather, prices, promotions, etc. In contrast to using a ``black-box'' machine learning model, Cyclic Boosting allows to interpret each individual prediction in terms of influence of the different features.

@@

To predict an individual PDF using a parametric approach, one has to rely on a model assumption about the underlying distribution of the random variable to be predicted. As discussed earlier, the NBD is well routed in theoretical arguments to model customer demand. Its parameters can be modeled by two independent models, one to estimate the mean and the other for the variance. At least in principle, any method can be used. However, as discussed in sec. \ref{sec:LitRev}, machine learning algorithms are ideally suited for the task of demand forecasting and in the following, we will use the Cyclic Boosting algorithm to benefit in particular from explainable decisions rather than black-box approaches. Furthermore, the regularization approach used during the training of the Cyclic Boosting algorithm allows a dedicated treatment of the underlying NBD model, which is another major benefit compared to a standard ``off-the-shelf'' machine learning algorithm. This means we use two subsequent Cyclic Boosting models in order to estimate the parameters of each individual PDF that we need to forecast. The first model is used to estimate the mean and the second to estimate the variance. The features may or may not differ between the mean and variance estimation models, and it can be beneficial to include the corresponding mean predictions as feature in the variance model. The assigned mean and variance predictions can then be used to generate individual PDFs using the parameterization of the NBD for each sample.

@@

Cyclic Boosting \cite{Wick2019} is a type of generalized additive model using a cyclic coordinate descent optimization and featuring a boosting-like update of parameters. Major benefits of Cyclic Boosting are its accuracy, performance, even at large scale, and providing fully explainable predictions, which are of vital importance in practical applications.

The main idea of this algorithm is the following: First, each feature, denoted by index $j$, is discretized appropriately into $k$ bins to reflect the specific behavior of the feature. The global mean $\mu$ is determined from all values $y$ of the  target variable $Y \in [0,\infty)$ observed in the data. Single data records, for example, the sales corresponding to a specific product-location-date combination along with all relevant features, are indexed by $i$.
The individual predictions $\hat{y_i}$  can then be calculated as:
\begin{equation} \label{eqn:cb}
\hat{y}_i = \mu \cdot \prod \limits_{j=1}^p f^k_j \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
\end{equation}
The factors $f^k_j$ are the model parameters that are determined iteratively from the features until the algorithm converges. During training, regularization techniques are applied to avoid overfitting and improve the generalization ability of the algorithm. The deviation of each factor from $f^k_j=1$ can then be used to explain how a specific feature contributes to each individual prediction.

@@

For a first qualitative assessment, we make use of the probability integral transform, see e.g. \cite{Angus1994,casella2002statistical}, which states that a random variable distributed according to the CDF of another random variable is uniformly distributed. We therefore expect that distribution of the corresponding actually observed CDF values of the individual PDF predictions is uniform if the predicted PDF is calibrated correctly, {\em regardless} of the shape of the predicted distribution. Any deviation can be interpreted as a hint that the predicted PDF is not fully correct \cite{diebold1998vevaluating}.

\noindent
The CDF of a PDF $f(x)$ is defined as:

@@

In case of discrete probability functions, such as the negative binomial function, the same argument still holds, but the definition of the inverse quantile function is replaced by the generalized inverse: $F^{-1}(y) = \mathrm{\inf \left \{x : F(x)>y\right  \} }$ for $y \in [0,1]$, see e.g. \cite[p. 54]{casella2002statistical}. In order to obtain a uniform distribution for discrete PDFs that is comparable to the case of continuous distributions, the histogram holding the values of the CDF is filled using random numbers according to the intervals of the CDF. For example, if the sales of zero items accounts for 75 percent of the observed sales distribution for this item, the value of the CDF function that is used to fill the histogram in case of zero actual sales is randomly chosen in the interval $[0, 0.75]$. Proceeding similarly for all other observed values, with the intervals from which to randomly choose values to fill in the histogram defined by the CDF values of the corresponding discrete sales value and the one below (e.g. for 3 actual sales: random pick between discrete CDF values for 2 and 3), the resulting histogram of CDF values is again uniform as in the case of a continuous PDF.

@@

Then we compare the actual observed sales value (corrected for censored data if necessary) to different quantiles of the corresponding predicted distribution for each data record and average over a larger data sample. For example, if we wanted to check that the median of the distribution, corresponding to the $0.5$ quantile, is predicted correctly by the machine learning model, we would compare the quantile value $0.5$ to the ratio of quantile values (again randomly chosen from the corresponding range of CDF values for discrete target values, as described in sec. \ref{sec:cdf_histo}) of observed sales records being lower/higher than $0.5$. In other words, in case of the median, 50 percent of the ex post observed target values should be observed below the median of the corresponding individual predicted PDF and 50 percent above.

In order to judge whether the overall shape of the predicted distributions is predicted correctly, we repeat this procedure for a range of quantiles, for example $q = 0.1, 0.2, \ldots, 0.9$. However, we are free to choose which quantiles to look at, and in specific situations it might be advisable to look at the tails of the distribution in more detail, to make sure that even relatively rare events are estimated correctly by the machine learning algorithm, and add more quantiles for comparison in the region between, say $q = 0.95$ and $q = 0.99$. In the following, we call this method {\em quantile profile plot}. Profile plots are akin to scatter plots and described in more detail in appendix \ref{sec:profile}.

@@

In the same way as the method of filling all observed CDF values of the individual PDF predictions in a histogram and compare to a uniform distribution (described in sec. \ref{sec:cdf_histo}), quantile profile plots do not work on individual PDF predictions but require a statistical population. However, calculating the fractions of the quantile profile plots globally, i.e., over all samples in the data set, might not reveal certain deficiencies for a subset, e.g., a specific store or product group in the example of retail demand forecasts. Therefore, we combine the approach described above with the method of profile plots, where the quantity on the $x$-axis of the quantile profile plot can be any variable of the data set at hand.

@@

\noindent
The different \texttt{item\_store} values for mean (higher than $1$) and variance model factors (lower than $1$), including the static one-dimensional features \texttt{store\_id} and \texttt{item\_id} and their two-dimensional combination, represent the fact that this product-location combination sells more than the average of all product-location combinations, and the dispersion parameter $1/r$ in \eqn \eqref{eqn:variance_r} in turn tends towards lower values. The explanation of the model for the peak in sales and mean prediction on February 7, 2016 is the event Super Bowl. The variance prediction for this day is driven by two competing factors:  the Super Bowl event itself drives  the  higher dispersion parameter higher, and the high mean prediction drives the lower dispersion parameter lower.

@@

\fig \ref{fig:invquant_mean} shows the quantile profile plot using the predicted mean as the $x$-axis of the graph, meaning that the distributions are grouped such that the 15 columns correspond to mean predictions in 12 intervals with widths of 5 from 0.0 to 60.0 and 3 remaining intervals $[60.0, 70.0]$, $(70.0, 80.0]$, and $(80.0, 100.0]$ (with mean predictions higher than 100.0 included in the highest interval), while we aggregate over all locations, items, and sales dates. The  quantile values are chosen to be the same as in \fig \ref{fig:invquant_dayofweek}. While the Poisson model again shows a significant pattern of too narrow PDF predictions across the full range of mean prediction values, we can also see several deviations from the expected uniform behavior for the negative binomial model. For mean predictions around $30$, the negative binomial model deviates significantly from the expected behavior, with the shape of the deviations pointing towards too broad PDF predictions (as also observed in the CDF histogram above). And for mean predictions higher than $80$, the pattern of the deviations indicate an underprediction of the mean parameter. The deviations at very low predicted mean values reflect the complications of slow-sellers described in the context of CDF histograms above.

@@

Once the full probability density functions are forecast, their evaluation poses significant challenges:
Common metrics, such as the mean absolute deviation or mean squared error, only take point estimates into account and are generally not suitable for the evaluation of predicted probability distributions. Furthermore, since the distributions are predicted for individual events, statistical techniques aiming at the comparison of distributions using many events cannot be used in general. Instead, novel techniques using both the histogram of the probability integral transformation as well as quantile profile histograms allow a detailed investigation into the behavior of the predictions. Finally, a quantitative assessment resulting in a single number can be obtained using metrics such as the earth mover distance in a comparison between the histogram of the probability integral transformation and the expected uniform distribution.

@@

In many cases, scatter plots are used to study the behavior of two distributions or sets of data points visually. However, even for moderate amount of data, this approach becomes quickly difficult. To illustrate this, a sample of $(x,y)$ data points was obtained in the following way: The distribution of $x$ values was obtained by generating 5,000 samples of Gaussian distributed random numbers $X \sim {\cal N}(0.0,2.0)$ and the $y$ values are obtained via $Y \sim X +  {\cal N}(2.0,1.5)$. \fig \ref{fig:scatter} shows the marginal distributions for $x$ and $y$ as well as a scatter plot of $x$ vs. $y$.
